---
title: "Create Custom Assertions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create_custom_assertions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(assertions)
```

## Getting Started

Do you have a custom assertion that you want to use repeatedly in your code? Then create your own assertion functions with the `assert_create()` function!.

Lets start by recreating the `assert_character()` assertion using `assert_create()`:

```{r, eval=FALSE}
# Create a function that asserts input is a character value
assert_character_2 <- assert_create(
  func = is.character(x), # Returns TRUE/FALSE for assertion PASS/FAIL
  default_error_msg = "'{arg_name}' must be a character value, not a {class(arg_value)}"
  )

# Assertion passes if input is a character value
assert_character_2("abc")

# But throws the expected error if input is not a character value
assert_character_2(123)

```

To create a custom assertion function, you need to supply two arguments to assert_create():

`func`: a **function** that take an object to assert and returns TRUE or FALSE depending on whether the assertion should **pass** or **fail**.

`default_error_msg`: a **character string** providing an **error message** in case the assertion **fails**. This string can include special termss such as:

1.  **{arg_name}** to refer to the name of the variable being checked
2.  **{arg_value}** to refer to the value of the variable.
3.  **{code_to_evaluate}** to evaluate code within the error message. Customise '*code_to_evaluate*'. e.g `{class(arg_name)}`
4.  **{.strong bold_text}** to perform inline formatting. Customise '*bold_text*'

## Advanced Assertions

Sometimes you need more than one error message per assertion. In this situation, often lots of individual assertions are still usually the best choice, but that's not always the case.

So you also have the option of using `{assert_create_advanced()}`.

This function works in a similar way to `assert_create{}` but it takes only one argument: `func`

**What does `func`** **expect?**

1.  **(like `assert_create`)** Function should return TRUE when assertion passes
2.  **(unlike `assert_create`)** Function should return a character string containing the error message text when assertion fails. This lets you implement different error messages for different ways in which the argument fails the assertion.
